---
title: "P8105 Homework 1"
author: "Xinzhuo Zhu"
date: "2026-09-25"
output: github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
# Problem 1

```{r}
data("penguins", package = "palmerpenguins")
```
The `penguins` dataset contains measurements and characteristics of penguins, including species, island, bill length, bill depth, flipper length, body mass, sex, and year. The species included in the dataset are `r paste(unique(penguins$species), collapse = ", ")`, and the islands represented are `r paste(unique(penguins$island), collapse = ", ")`.

The dataset contains `r nrow(penguins)` rows and `r ncol(penguins)` columns. The mean flipper length is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2)` mm.

```{r, warning=FALSE, message=FALSE}
penguin_plot =
  ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point()

penguin_plot

ggsave("penguin_scatterplot.png", plot = penguin_plot)
```
# Problem 2

```{r}
set.seed(1234)

problem2_df =
  tibble(
    random_sample = rnorm(10),
    positive = random_sample > 0,
    character_vector = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    factor_vector = factor(c("low", "medium", "high", "low", "medium",
                             "high", "low", "medium", "high", "low"))
  )

problem2_df
```
```{r, warning=FALSE}
mean(pull(problem2_df, random_sample))
mean(pull(problem2_df, positive))
mean(pull(problem2_df, character_vector))
mean(pull(problem2_df, factor_vector))
```

The mean can be calculated for the numeric and logical variables, but not for the character or factor variables. For the logical variable, R treats `TRUE` as 1 and `FALSE` as 0, so the mean represents the proportion of values that are `TRUE`. The character and factor variables return `NA` because they are not numeric or logical.
```{r, results='hide', warning=FALSE}
as.numeric(pull(problem2_df, positive))
as.numeric(pull(problem2_df, character_vector))
as.numeric(pull(problem2_df, factor_vector))
```
Converting the logical variable to numeric changes `FALSE` to 0 and `TRUE` to 1, which explains why its mean can be calculated. Converting the character variable to numeric produces `NA` values because the letters cannot be interpreted as numbers. Converting the factor variable to numeric returns the integer codes corresponding to its factor levels rather than meaningful numeric values. This helps explain why the logical variable can be averaged directly, while the character and factor variables cannot.
